THESIS PROJECT PROPOSAL

### 2019-11-01 project_ashish/paper.tex
Update paper.tex

@@

The Standard Model (SM) of particle physics has successfully described most of the experimental data till now but a very large number of
free-parameters like fine structure constant \alpha, Weinberg angle or weak mixing angle \theta_W, the coupling constant of strong 
interaction, electroweak symmetry breaking energy scale, Higgs potential self coupling or the Higgs mass, three mixing angles and the 
CP-violating phase of the CKM matrix which tells us how quarks of different color charge mix with each other, nine yukawa couplings 
which determine the mass of nine charged fermions and fine tunings related to the origin of masses robustly suggests new physics beyond 
the SM (BSM).There are lot of specific beyond BSM theories and most of these models involve heavy fields. In order to identify which new 
physics lies beyond the electroweak (EW) scale, the new parameters of such theories may be constrained by the actual, low energy, 
experiments. This approach requires studying each model individually, and calculating every possible observable. There is another 
approach immitating Fermi’s treatment of beta decay which consists of considering the SM as the first order approximation of the actual 
theory and by completing it with a series of higher dimensional operators. When the electroweak symmetry breaking takes place well below 
the mass of the new particles, the BSM physics is taken into account at the EW scale and below by adding higher dimensional operators to 
the SM Lagrangian. They are built out of SM fields and supposed to be invariant under its gauge group. Those operators are the low-
energy remnant of the high energy theory. This approach does not pretend to guess the complete high-energy model and it is based solely 
on the symmetry of the  theory. The operators are general and the only model dependence is encoded in the size of the operator 
coefficients, which is to be set from experiments.

>>>>

Until the 90's the existence of almost all the particles of the SM were confirmed except the top quark and the Higgs boson. These had 
eluded previous experiments due to difficulties in the production or reconstruction of its decays. The top quark was discovered in 1995 
in the Tevatrón collider of the Fermilab laboratory, this proton collider operated with an energy of 1.8 TeV at the center of mass up to 
2010. The LHC collider at the CERN laboratory in Geneva, Switzerland, began its operations in 2010 colliding protons at 7 TeV and then 
at 8 and 13 TeV.

The Compact Muon Solenoid (CMS) is based on the Large Hadron Collider (LHC) located at CERN. It is designed to detect particles known as 
muons very accurately. The CMS detector has the form of a cylindrical onion, with several concentric layers of components. Needed
a powerful magnet to bend charged particles as they move away from the point of collision to identify the charge of the particles to 
bend them in opposite directions and measure your momentum. A silicon tracker, made of about 75 million electronic sensors individual arranged in concentric layers, identifies the routes taken by these charged particles bent with very high precision. The muons are not detected by electromagnetic or hadron calorimeters, so special sub-detectors are outside to detect them once they have crossed the solenoid. In 2012, the ATLAS and CMS collaborations, with detectors at two points where the protons collide in the LHC, announced the discovery of a new boson with a mass of 125 GeV. So far, all measurements of the properties of this boson are consistent with those of the Higgs boson of the standard model (SM).

@@

The Standard Model (SM) of particle physics has successfully described most of the experimental data till now but a very large number of
free-parameters like fine structure constant \alpha, Weinberg angle or weak mixing angle \theta_W, the coupling constant of strong 
interaction, electroweak symmetry breaking energy scale, Higgs potential self coupling or the Higgs mass, three mixing angles and the 
CP-violating phase of the CKM matrix which tells us how quarks of different color charge mix with each other, nine yukawa couplings 
which determine the mass of nine charged fermions and fine tunings related to the origin of masses robustly suggests new physics beyond 
the SM (BSM).There are lot of specific beyond BSM theories and most of these models involve heavy fields. In order to identify which new 
physics lies beyond the electroweak (EW) scale, the new parameters of such theories may be constrained by the actual, low energy, 
experiments. This approach requires studying each model individually, and calculating every possible observable. There is another 
approach immitating Fermi’s treatment of beta decay which consists of considering the SM as the first order approximation of the actual 
theory and by completing it with a series of higher dimensional operators. When the electroweak symmetry breaking takes place well below 
the mass of the new particles, the BSM physics is taken into account at the EW scale and below by adding higher dimensional operators to 
the SM Lagrangian. They are built out of SM fields and supposed to be invariant under its gauge group. Those operators are the low-
energy residue of the high energy theory. This approach does not pretend to guess the complete high-energy model and it is based solely 
on the symmetry of the  theory. The operators are general and the only model dependence is encoded in the size of the operator 
coefficients, which is to be set from experiments.
